---
title: "PMMD Experimental Medicine simulation"
output: html
---

# Libraries

```{r}

library(tidyverse)
library(ggplot2)
library(lattice)
library(wesanderson)

pal <- wes_palette("Zissou1", 5, type = "continuous")

```

# Simulation

Example
- compare: clinical vs mood (self-report weekly) vs wearables

Aiming to recover:
- time to detection of depression

Simulation set-up:
- similar duration of episode between people
- variable amplitude between people
- similar lag between modalities across individuals

Assumptions:
- that lag between modalities across individuals is similar (+/- some error)

## Sine wave

Additional assumptions:
- sine wave (better if could use function that stops oscillating after returning to 0)

### Simulate physical activity (daily)

```{r}

# Simulate physical activity level (daily)
pday <- seq(from = 1, to = 364, by = 1) # number of days of observation
duration <- 300
b <- 2*pi/duration
n <- 20 # sample size

pp <- rnorm(n, b, 0.001) # different period for each ind
ps <- rnorm(n, duration/4, 0.01) # different shift of episode for each ind
pa <- rnorm(n, 1, 0.5)

phys <- matrix(nrow = length(pday), ncol = n)
for (i in 1:ncol(phys)) {
 phys[,i] <- pa[i]*sin(pp[i]*(pday + ps[i])) + rnorm(length(pday), 0, 0.3) # add noise
}
phys.df <- data.frame(day = pday, phys, type = "phys") %>% pivot_longer(cols = (-c(day, type)), names_to = "ind", values_to = "polarity")

ggplot(phys.df, aes(x = day, y = polarity)) + geom_point() +
  facet_wrap(~ ind)

```

## Simulate mood (weekly)

```{r}

# Simulate mood diary (weekly)
mday <- seq(from = 1, to = 364, by = 7) # number of days of observation
duration <- 300
b <- 2*pi/duration
n <- 20 # sample size
lag <- -7

# Assume that the relationship between activity and mood for each individual has 
# - similar period
# - similar lag (specified as 7 days here)
mp <- NULL
ms <- NULL
ma <- NULL
for (i in 1:n) {
  mp <- c(mp, rnorm(1, pp[i], 0.0001))
  ms <- c(ms, rnorm(1, ps[i] + lag, 0.0001))
  ma <- c(ma, rnorm(1, pa[i], 0.0001))
}

mood <- matrix(nrow = length(mday), ncol = n)
for (i in 1:ncol(mood)) {
 mood[,i] <- ma[i]*sin(mp[i]*(mday + ms[i])) + rnorm(length(mday), 0, 0.3) # add noise
}
mood.df <- data.frame(day = mday, mood, type = "mood") %>% pivot_longer(cols = (-c(day, type)), names_to = "ind", values_to = "polarity")

ggplot(mood.df, aes(x = day, y = polarity)) + geom_point() +
  facet_wrap(~ ind)

```

## Clinical (monthly)

```{r}

# Simulate clinical (monthly)
cday <- seq(from = 1, to = 364, by = 28) # number of days of observation
duration <- 300
b <- 2*pi/duration
n <- 20 # sample size
lag <- -14

# Assume that the relationship between activity and clinical detection for each individual has 
# - similar period
# - similar lag (specified as 14 days here)
cp <- NULL
cs <- NULL
ca <- NULL
for (i in 1:n) {
  cp <- c(cp, rnorm(1, pp[i], 0.0001))
  cs <- c(cs, rnorm(1, ps[i] + lag, 0.0001))
  ca <- c(ca, rnorm(1, pa[i], 0.0001))
}

clin <- matrix(nrow = length(cday), ncol = n)
for (i in 1:ncol(clin)) {
 clin[,i] <- ca[i]*sin(cp[i]*(cday + cs[i])) + rnorm(length(cday), 0, 0.3) # add noise
}
clin.df <- data.frame(day = cday, clin, type = "clinical") %>% pivot_longer(cols = (-c(day, type)), names_to = "ind", values_to = "polarity")

ggplot(clin.df, aes(x = day, y = polarity)) + geom_point() +
  facet_wrap(~ ind)

```

## Visualise all data together

```{r}

sine.df <- full_join(phys.df, mood.df) %>% full_join(., clin.df)
ggplot(sine.df, aes(x = day, y = polarity, colour = type)) + 
  geom_point(alpha = 0.5) + 
  scale_colour_manual(values = pal) +
  theme_bw() + theme(legend.position = "none") + 
  facet_wrap(~ ind)

```

## Rational 

library(phonTools) # for sinc() function
library(splines)
library(broom)
library(ggformula)

### Simulate physical activity (daily)

```{r}



```

# Analysis

## Spline

```{r}

spline()
foo <- sine.df %>% filter(ind == "X1" & type == "phys")

tmp <- ggplot(sine.df, aes(x = day, y = polarity, colour = type)) + 
  geom_point(alpha = 0.) + 
  geom_spline() + 
  scale_colour_manual(values = pal) +
  theme_bw() +
  facet_wrap(~ ind)
tmp

tmp <- ggplot(sine.df, aes(x = day, y = polarity, colour = type)) + 
  geom_point(alpha = 0.5) + 
  geom_smooth(method = lm, 
              formula = y ~ splines::bs(x, knots = 2)) +
  scale_colour_manual(values = pal) +
  theme_bw() +
  facet_wrap(~ ind)
tmp

```

# Extensions

- changing midline: https://math.libretexts.org/Bookshelves/Precalculus/Book%3A_Precalculus__An_Investigation_of_Functions_%28Lippman_and_Rasmussen%29/07%3A_Trigonometric_Equations_and_Identities/7.04%3A_Modeling_Changing_Amplitude_and_Midline 
- dampened sine wave

```{r}

```