---
title: "PMMD BiMODAL simulation"
author: "Chloe Yap"
date: "`r Sys.Date()`"
output:
  html_document:
      toc: true
      number_sections: true
      toc_float: true
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Simulation set-up 

- compare: clinical (monthly) vs mood (self-report weekly) vs wearables (continuous but say daily)

Questions of interest for the human study:

- can wearables data predict onset of depression defined using mood self-report and/or clinical impression?
- how much earlier can wearables data predict depression (vs mood self-report and clinical impression)?

Questions for this simulation:

1. Can we recover simulated lag time between wearables vs mood; wearables vs clinical?
2. Can we identify some threshold when wearables data from the individual can definitively diagnose "depression" vs a mood-based definition of depression (eg. 2 consecutive negative points taken 1 week apart) vs clinical-based definition of depression (eg. 2 consecutive negative points taken 1 month apart)

Simulation set-up:

- thresholds for mood: y > 0.5 = elevated mood vs y < -0.5 = depressed mood
- similar duration of episode between people
- variable amplitude between people
- similar lag between modalities across individuals
  + lag for mood vs wearable = 7 days behind
  + lag for clinical vs wearable = 14 days behind

Issues:

- assuming in this simulation that lag between modalities across individuals is similar (+/- some error). This may not be true in practice!!
- assuming that the lag "looks the same" across modalities
- **ground truth mood might be trickier to assess than on first glance ... People who are manic often feel great. Also not necessarily able to assume that people are euthymic from the start of study collection. So what's ground truth baseline/euthymia?**

# Libraries

```{r, message=F}

library(rmarkdown)
library(tidyverse)
library(ggplot2)
library(lattice)
library(wesanderson)
library(ggformula)
library(lmtest)
library(gtsummary)

pal <- wes_palette("Zissou1", 5, type = "continuous")

set.seed(1234)

```

# Simulation for Aim 1

## Model as sine wave

- sine wave BUT better if could use function that stops oscillating after returning to 0 ... Assuming that people have mania --> depression --> euthymia afterwards

### Simulate wearables (daily sampling)

```{r}

# Simulate physical activity level (daily)
pday <- seq(from = 1, to = 364, by = 1) # number of days of observation
duration <- 300
b <- 2*pi/duration
n <- 20 # sample size

pp <- rnorm(n, b, 0.001) # different period for each ind
ps <- rnorm(n, duration/4, 0.01) # different shift of episode for each ind
pa <- rnorm(n, 1, 0.5)

phys <- matrix(nrow = length(pday), ncol = n)
for (i in 1:ncol(phys)) {
 phys[,i] <- pa[i]*sin(pp[i]*(pday + ps[i])) + rnorm(length(pday), 0, 0.3) # add noise
}
phys.df <- data.frame(day = pday, phys, type = "phys") %>% pivot_longer(cols = (-c(day, type)), names_to = "ind", values_to = "polarity")

ggplot(phys.df, aes(x = day, y = polarity)) + geom_point() +
  facet_wrap(~ ind) +
  geom_hline(yintercept=-0.5) + geom_hline(yintercept=0.5)

```

### Simulate mood (weekly)

```{r}

# Simulate mood diary (weekly)
mday <- seq(from = 1, to = 364, by = 7) # number of days of observation
duration <- 300
b <- 2*pi/duration
n <- 20 # sample size
lag <- -7

# Assume that the relationship between activity and mood for each individual has 
# - similar period
# - similar lag (specified as 7 days here)
mp <- NULL
ms <- NULL
ma <- NULL
for (i in 1:n) {
  mp <- c(mp, rnorm(1, pp[i], 0.0001))
  ms <- c(ms, rnorm(1, ps[i] + lag, 0.0001))
  ma <- c(ma, rnorm(1, pa[i], 0.0001))
}

mood <- matrix(nrow = length(mday), ncol = n)
for (i in 1:ncol(mood)) {
 mood[,i] <- ma[i]*sin(mp[i]*(mday + ms[i])) + rnorm(length(mday), 0, 0.3) # add noise
}
mood.df <- data.frame(day = mday, mood, type = "mood") %>% pivot_longer(cols = (-c(day, type)), names_to = "ind", values_to = "polarity")

ggplot(mood.df, aes(x = day, y = polarity)) + geom_point() +
  facet_wrap(~ ind) +
  geom_hline(yintercept=-0.5) + geom_hline(yintercept=0.5)

```

### Clinical (monthly)

```{r}

# Simulate clinical (monthly)
cday <- seq(from = 1, to = 364, by = 28) # number of days of observation
duration <- 300
b <- 2*pi/duration
n <- 20 # sample size
lag <- -14

# Assume that the relationship between activity and clinical detection for each individual has 
# - similar period
# - similar lag (specified as 14 days here)
cp <- NULL
cs <- NULL
ca <- NULL
for (i in 1:n) {
  cp <- c(cp, rnorm(1, pp[i], 0.0001))
  cs <- c(cs, rnorm(1, ps[i] + lag, 0.0001))
  ca <- c(ca, rnorm(1, pa[i], 0.0001))
}

clin <- matrix(nrow = length(cday), ncol = n)
for (i in 1:ncol(clin)) {
 clin[,i] <- ca[i]*sin(cp[i]*(cday + cs[i])) + rnorm(length(cday), 0, 0.3) # add noise
}
clin.df <- data.frame(day = cday, clin, type = "clinical") %>% pivot_longer(cols = (-c(day, type)), names_to = "ind", values_to = "polarity")

ggplot(clin.df, aes(x = day, y = polarity)) + geom_point() +
  facet_wrap(~ ind) +
  geom_hline(yintercept=-0.5) + geom_hline(yintercept=0.5)

```

### Visualise all data together

```{r}

sine.df <- full_join(phys.df, mood.df) %>% full_join(., clin.df)
ggplot(sine.df, aes(x = day, y = polarity, colour = type)) + 
  geom_point(alpha = 0.5) + 
  scale_colour_manual(values = pal) +
  theme_bw() + theme(legend.position = "none") + 
  facet_wrap(~ ind)

```

## Extensions to consider

library(phonTools) # for sinc() function
library(splines)
library(broom)
library(ggformula)

# Analysis

## Granger causality test

Questions:

- How to determine good number of lags?
- How to apply over many people?

To read in more depth: https://davegiles.blogspot.com/2011/04/testing-for-granger-causality.html

```{r, warnings = F, messages = F}

gt.moodphys <- list()
gt.physmood <- list()

for (i in 1:n) {
  tmp <- rbind(phys.df %>% filter(ind == paste("X", i, sep = "")),
             mood.df %>% filter(ind == paste("X", i, sep = ""))) %>%
      pivot_wider(id_cols = "day", names_from = "type", values_from = "polarity")

  paged_table(tmp)

  gt.moodphys[[i]] <- grangertest(mood ~ phys, order = 3, data = tmp)
  gt.physmood[[i]] <- grangertest(phys ~ mood, order = 3, data = tmp)

}

gt.moodphys.tbl <- lapply(gt.moodphys, tbl_regression)
gt.physmood.tbl <- lapply(gt.physmood, tbl_regression)

tbl_merge(tbls = gt.moodphys.tbl, tab_spanner = paste("X", seq(1, n, 1), sep = ""))
tbl_merge(tbls = gt.physmood.tbl, tab_spanner = paste("X", seq(1, n, 1), sep = ""))

```

### Plot p-values

False positive/negatives shown with:

- dotted line for p<0.05
- dashed line for Bonferroni correction
- Q1: true + for mood ~ phys / true - for phys ~ mood
- Q2: true + for mood ~ phys / false + for phys ~ mood
- Q3: false - for mood ~ phys / true - for phys ~ mood
- Q4: false - for mood ~ phys / false + for phys ~ mood

```{r}

gt.moodphys.p <- unlist(lapply(1:n, function(x) gt.moodphys[[x]]$`Pr(>F)`[2]))
gt.physmood.p <- unlist(lapply(1:n, function(x) gt.physmood[[x]]$`Pr(>F)`[2]))

gt.p.df <- data.frame(id = paste("X", seq(1, n, 1), sep = ""), gt_p_moodphys = gt.moodphys.p, gt_p_physmood = gt.physmood.p)

ggplot(gt.p.df, aes(x = -log10(gt_p_physmood), y = -log10(gt_p_moodphys))) +
  geom_point() +
  theme_bw() +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted") +
  geom_hline(yintercept = -log10(0.05/n), linetype = "dashed") +
  geom_vline(xintercept = -log10(0.05), linetype = "dotted") +
  geom_vline(xintercept = -log10(0.05/n), linetype = "dashed")

```

## Spline? Sine wave?

Not sure what I should use to flexibly fit a function (assuming that's the right thing to do) ...

Maybe try(?):

- fit some ?spline function for the monthly clinical data ("ground truth A")
- fit some ?spline function for the weekly mood diary data ("ground truth B")
- try and fit functions A and B with changeable left-shift for the wearables data and see if it picks up the lag time?

Limitations of this approach:

- don't know in real data if the exact same trend is followed across datasets! 
- might just be looking for any change preceding depression defined using ground truth A and ground truth B ... ?Bayesian approach since prior is strong and n is small?

```{r}

# spline() ???

foo <- sine.df %>% filter(ind == "X1" & type == "phys")

tmp <- ggplot(sine.df, aes(x = day, y = polarity, colour = type)) + 
  geom_point(alpha = 0.) + 
  geom_spline() + 
  scale_colour_manual(values = pal) +
  theme_bw() +
  facet_wrap(~ ind)
tmp

tmp <- ggplot(sine.df, aes(x = day, y = polarity, colour = type)) + 
  geom_point(alpha = 0.5) + 
  geom_smooth(method = lm, 
              formula = y ~ splines::bs(x, knots = 2)) +
  scale_colour_manual(values = pal) +
  theme_bw() +
  facet_wrap(~ ind)
tmp

```

# (IGNORE): Extensions

- changing midline: https://math.libretexts.org/Bookshelves/Precalculus/Book%3A_Precalculus__An_Investigation_of_Functions_%28Lippman_and_Rasmussen%29/07%3A_Trigonometric_Equations_and_Identities/7.04%3A_Modeling_Changing_Amplitude_and_Midline 
- dampened sine wave

```{r}

```